Combinatorics of the geometry of Wilson loop diagrams

### 2019-03-22 General WLD combinatorics.tex
finished with the dimension proof!

Recall that $I_1$ determines the shape of the Le diagram. By Lemma~\ref{lem I} we know that ${n,n-1,n-2, n-3 \not\in I_1^{V}}$ (which yields the leftmost four columns of the diagram for $V$) and that $I_1^{W} = I_1^{V}\cup \{n-3\}$. This implies that the right hand boundary of the shape of $V$ is the same as the right hand boundary of the shape of $W$ except that $W$ has one additional row of 3 boxes while $V$ has an additional column in the $n-3$ position; that is, an extra column fourth from the left.
\end{proof}

@@

Then $I_{n-2}^{W}$ and $I_{n-3}^{W}$ contribute the following plusses to the Le diagram of $W$:
  \begin{itemize}
  \item All of the $I_{n-1}^{V}$ plusses.
  \item An $(n-3)\rightarrow (n-2)$ plus and a $(n-1)\rightarrow (n-1)$ plus.
  \item In the Le diagram of $V$, $I_{n-2}^{V}$ contributes an $n-5\rightarrow n-2$ plus and no other term in the Grassmann necklace of $V$ gives a plus in this column.  $I_{n-2}^W$ does not contribute this $+$ but yields an $(n-5)\rightarrow (n-1)$ plus instead.
  \item All other plusses of $I_{n-2}^{V}$.
  \item In the Le diagram of $V$, $I_{n-3}^{V}$ gives a plus in the $n-3$ column.  This $+$ is shifted one square to the left (into the $n-2$ column) in the Le diagram of $W$.
  \item All other plusses of $I_{n-3}^{V}$.
  \end{itemize}
Furthermore, no element of the Grassmann necklace of $V$ contributes a $(n-5)\rightarrow (n-1)$ plus.
\end{lem}

\begin{proof}
  By Lemma~\ref{lem I} we have $I_{n-2}^{W}= (I_{n}^{V} \setminus \{n-5\})\cup \{n-1,n-2\}$. We are necessarily in the case where $W$ admits a propagator $q$ (Case 2 of Figure~\ref{fig special p}) and we can see from the location of $q$ that $n-2\not\in I_{n-3}^{V}$ and that $n-5$ is the largest row index in $\mathcal{B}$.  Thus this section of the Gramann necklace of $V$ looks like
  \begin{equation}\label{eq necklace}
  I_{n-3}^{V} \underset{\substack{n-3\text{ out}\\n-2\text{ in}}}{\rightarrow} I_{n-2}^{V} \underset{\substack{n-2\text{ out}\\n-5\text{ in}}}{\rightarrow} I_{n-1}^{V} \underset{\substack{n-1\text{ out}\\\text{something in}}}{\rightarrow} I_{n}^{V} \underset{\substack{n \text{ out}\\\text{something in}}}{\rightarrow} I_1^{V}
  \end{equation}
  where the first ``something'' is either $n$ or an element of $I_1^{V}$ and the second ``something'' is an element of $I_1^{V}$.  %Additionally all elements not explicitly mentioned must be in $I_1^{V}$ as they remain unchanged through this portion of the necklace.

  Using this information we can determine the symmetric difference of $I_{n-2}^{V}$ and $I_1^{V}$, which has size $\leq 3$ by the definition of a Grassmann necklace. We have $n-1, n-2 \in I_{n-2}^{V} \setminus I_1^V$ for certain, and the third element (if it exists) must be $n$. On the other hand, we have $n-5 \in I_1^{V} \setminus I_{n-2}^V$, along with at least one and at most two other elements.  If there is one such element call it $a$.  If there are two call them $a$ and $b$ with $a>b$.

  This means that the plusses in the Le diagram of $V$ coming from $I_{n-2}^{V}$ are as in the first part of Figure~\ref{fig messy}.  Stepping to $I_{n-1}^{V}$ simply removes the $n-5\rightarrow n-2$ path (equation \eqref{eq necklace}); this is illustrated in the second part of Figure~\ref{fig messy}.

  \begin{figure}
    \includegraphics{messy}
    \caption{Plusses coming from $I_{n-2}^{V}$ (top left),  $I_{n-1}^{V}$ (top right) and $I_{n}^{V}$ (bottom), in the case where $n-1, n-2\in I_{n-2}^{V}$.  The blue lines are the non-intersecting paths defining the position of the plusses.  The dashed blue lines may or may not appear, but if one appears then they both do.}\label{fig messy}
  \end{figure}

  Stepping to $I_{n}^{V}$, $n-1$ is removed and either $n$ is put in if it was not there before, or one of $a$ or $b$ is put in and hence is no longer available as the starting point for a path.  This gives two possible configurations illustrated in the bottom two parts of Figure~\ref{fig messy}.

@@

  Now consider $I_{n-3}^{V}$.  By \eqref{eq necklace} $I_{n-3}^{V}$ contributes almost the same plusses as $I_{n_2}^{V}$: the only difference is that it contributes an $(n-5)\rightarrow (n-3)$ plus in place of the $(n-5)\rightarrow (n-2)$ plus.  Since we must have $n-3\in I_{n-3}^{V}$ (because propagator $q$ exists in $V$), it follows from Lemma\ref{lem I} that $I_{n-3}^{W} = I_{n-3}^{V}\cup\{n-2\}$.  Thus the paths for $I^{W}_{n-3}$ are the same as for $I^{V}_{n-3}$ except that the path that did go to $n-3$ now goes to $n-2$.  This cannot conflict with another path since \eqref{eq necklace} shows that $n-2$ only appears in $I_{n-2}^{V}$ among the necklace elements of $V$.

  Also note that $I_{n-3}^{V}, I_{n-2}^{V}$, and $I_{n-1}^{V}$ share their plusses outside of the $n-3$ and $n-2$ columns.  This proves the remaining bullet points of the lemma, leaving only the claim about the $(n-5)\rightarrow (n-1)$ plus.

  So suppose there were a $n-5\rightarrow n-1$ plus in the Le diagram of $V$.  By Algorithm~\ref{alg:GN to Le}, this plus could only be contributed by a necklace term that satisfies $n-5 \not\in I_*^{V}$.  By \eqref{eq necklace}, the only terms that could have this property are $I_{n-4}^{V}$, $I_{n-3}^{V}$, and $I_{n-2}^{V}$. The analysis above shows that it does not come from $I_{n-2}^V$ or $I_{n-3}^V$.

  Recall that $n-5$ is the largest row index in the Le diagram of $V$, and $n-5 \in I_1^V \setminus I_{n-4}^V$ so we do have a path starting at $n-5$. Since $n-4 \in I_{n-4}^V$ by the location of the propagator $q$, we must have a path ending at $n-4$. Since the paths cannot cross, this implies that $_{n-4}^V$ must contribute a $(n-5) \rightarrow (n-4)$ plus, and hence cannot contribute a $(n-5)\rightarrow (n-1)$ plus.

@@

>>>>

Recall that $I_1$ determines the shape of the Le diagram. By Lemma~\ref{lem I} we know that
\newline ${n,n-1,n-2, n-3 \not\in I_1^{V}}$ (which yields the leftmost four columns of the diagram for $V$) and that $I_1^{W} = I_1^{V}\cup \{n-3\}$. This implies that the right hand boundary of the shape of $V$ is the same as the right hand boundary of the shape of $W$ except that $W$ has one additional row of 3 boxes while $V$ has an additional column in the $n-3$ position; that is, an extra column fourth from the left.
\end{proof}

@@

Then $I_{n-2}^{W}$ and $I_{n-3}^{W}$ contribute the following plusses to the Le diagram of $W$
  \begin{itemize}
  \item An $(n-3)\rightarrow (n-2)$ plus and an $(n-5)\rightarrow (n-1)$ plus.
  \item All of the $I_{n-1}^{V}$ plusses.
  \item In the Le diagram of $V$, $I_{n-2}^{V}$ contributes an $n-5\rightarrow n-2$ plus and no other term in the Grassmann necklace of $V$ gives a plus in this column.  $I_{n-2}^W$ does not contribute this $+$ but yields an $n-5\rightarrow n-1$ plus instead.
  \item All other plusses of $I_{n-2}^{V}$.
  \item In the Le diagram of $V$, $I_{n-3}^{V}$ gives a plus in the $n-3$ column.  This $+$ is shifted over into the $n-2$ column in $W$.
  \item All other plusses of $I_{n-3}^{V}$.
  \end{itemize}
Furthermore, no element of the Gramann necklace of $V$ gives an $n-5\rightarrow n-1$ plus.
\end{lem}

\begin{proof}
  By Lemma~\ref{lem I} $I_{n-2}^{W}= (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\}$.  Also, by the location of $q$ in the WLD,  $n-2\not\in I_{n-3}^{V}$ and and $n-5$ is the index of the lowest vertical edge in $\mathcal{B}$.  Thus this section of the Gramann necklace of $V$ looks like
  \begin{equation}\label{eq necklace}
  I_{n-3}^{V} \underset{\substack{n-3\text{ out}\\n-2\text{ in}}}{\rightarrow} I_{n-2}^{V} \underset{\substack{n-2\text{ out}\\n-5\text{ in}}}{\rightarrow} I_{n-1}^{V} \underset{\substack{n-1\text{ out}\\\text{something in}}}{\rightarrow} I_{n}^{V} \underset{\substack{n \text{ out}\\\text{something in}}}{\rightarrow} I_1^{V}
  \end{equation}
  where the first ``something'' is either $n$ or an element of $I_1^{V}$ and the second ``something'' is an element of $I_1^{V}$.  Additionally all elements not explicitly mentioned must be in $I_1^{V}$ as they remain unchanged through this portion of the necklace.

  Using this information now determine the symmetric difference of $I_{n-2}^{V}$ and $I_1^{V}$: $n-1, n-2$ and possibly $n$ are in $I_{n-2}^{V}$ but not in $I_1^{V}$.  $n-5$ is in $I_1^{V}$ as are at least one and at most two other elements.  If there is one such element call it $a$.  If there are two call them $a$ and $b$ with $a>b$.  This means that the plusses in the Le diagram of $V$ coming from $I_{n-2}^{V}$ are as in the first part of Figure~\ref{fig messy}.  Stepping to $I_{n-1}^{V}$ simply removes the $n-5\rightarrow n-2$ path, see the second part of Figure~\ref{fig messy}.

  \begin{figure}
    \includegraphics{messy}
    \caption{Plusses coming from $I_{n-2}^{V}$ (top left),  $I_{n-1}^{V}$ (top right) and $I_{n}^{V}$ bottom when $n-1, n-2\in I_{n-1}^{V}$.  The blue lines are the non-intersecting paths.  The dashed blue lines may or may not appear, but if one appears then they both do.}\label{fig messy}
  \end{figure}

  Stepping to $I_{n}^{V}$, $n-1$ is taken out and either $n$ is put in if it was not there before, or one of $a$ or $b$ is put in and hence no longer available as a right end for a path.  This gives two possible configurations illustrated in the bottom two parts of Figure~\ref{fig messy}.

@@

  Now consider $I_{n-3}^{V}$.  By \eqref{eq necklace} $I_{n-3}^{V}$ contributes the same plusses as $I_{n_2}^{V}$ except that it contributes an $n-5\rightarrow n-3$ plus in place of the $n-5\rightarrow n-2$ plus.  Also, we have $n-3\in I_{n-3}^{V}$ be the location of $q$ and so $I_{k}^{W} = I_k^{V}\cup\{n-2\}$.  Thus the paths for $I^{W}_{n-3}$ are the same as for $I^{V}_{n-3}$ except that the path that did go to $n-3$ now goes to $n-2$.  This cannot conflict with another path since \eqref{eq necklace} shows that $n-2$ only appears in $I_{n-2}^{V}$ among the necklace elements of $V$.

  Also note that $I_{n-3}^{V}, I_{n-2}^{V}$, and $I_{n-1}^{V}$ share their plusses outside of the $n-3$ and $n-2$ columns.  This proves the remaining statements of the lemma except the furthermore.

  Finally, suppose there were a $n-5\rightarrow n-1$ plus in the Le diagram of $V$.  By the algorithm, it would have to come when $n-5 \not\in I_j^{V}$.  By \eqref{eq necklace} this means that it would have to come from $I_{n-4}^{V}$, $I_{n-3}^{V}$, or $I_{n-2}^{V}$. The analysis above shows it does not come from $I_{n-3}^{V}$ or $I_{n-2}^{V}$.   Now, $n-4\in I_{n-4}^{V}$ by the location of $q$ and so $I_{n-4}^{V}$ must give an $n-5\rightarrow n-4$ plus and so cannot give an $n-5\rightarrow n-1$ plus.

@@

\note{This is a reorganised version of the previous lemma, keeping both in place for now for comparison.}
\begin{lem}\label{lem n-2 bad v2}
Let $V$ and $W$ be as in Lemma~\ref{lem I}, and suppose that $n-2,n-1 \in I_{n-2}^{V}$. Then:
  \begin{enumerate}
  \item $I_{n-2}^W$ contributes an $(n-3)\rightarrow (n-2)$ plus and an $(n-5)\rightarrow (n-1)$ plus, along with any plusses contributed by $I_n^V$.
  \item $I_{n-3}^W$ contributes the same plusses as $I_{n-3}^V$, except that the plus in the $n-3$ column is shifted one square left into the $n-2$ column.
  \item In the Le diagram of $V$, $I_{n-2}^{V}$ contributes an $(n-5)\rightarrow (n-2)$ plus and no other term in the Grassmann necklace of $V$ gives a plus in this column.  $I_{n-2}^W$ does not contribute this $+$ but yields the $(n-5)\rightarrow (n-1)$ plus from point (1) instead.
  \item All other plusses from $I_{n-2}^{V}$ and all plusses from $I_{n-1}^V$ were already contributed by $I_{n-3}^V$, so they yield no new information.
  \item No element of the Grassmann necklace of $V$ contributes a $(n-5)\rightarrow (n-1)$ plus.
  \end{enumerate} 
\end{lem}

\begin{proof}
  By Lemma~\ref{lem I} we have $I_{n-2}^{W}= (I_{n}^{V} \setminus \{n-5\})\cup \{n-1,n-2\}$ and $I_1^W = I_1^V\cup \{n-3\}$. We are necessarily in the case where $W$ admits a propagator $q$ (the right hand case of Figure~\ref{fig special p}) and we can see from the location of $q$ that $n-2\not\in I_{n-3}^{V}$ and that $n-5$ is the index of the lowest vertical edge in $\mathcal{B}$.  Thus this section of the Gramann necklace of $V$ looks like

  \begin{equation}\label{eq necklace}
  I_{n-3}^{V} \xrightarrow[\substack{n-3\text{ out}\\n-2\text{ in}}]{} I_{n-2}^{V} \xrightarrow[\substack{n-2\text{ out}\\n-5\text{ in}}]{} I_{n-1}^{V} \xrightarrow[\substack{n-1\text{ out}\\\text{something in}}]{} I_{n}^{V}  \xrightarrow[\substack{n \text{ out}\\\text{something in}}]{\text{\ no change, {\em or}\ }} I_1^{V}
  \end{equation}
  where the first ``something'' is either $n$ or an element of $I_1^{V}$, and the second ``something'' (if it exists) is an element of $I_1^{V}$. If $n\not\in I_n^V$ then $I_n^V = I_1^V$.

  Using this information we can determine the symmetric difference of $I_{n-2}^{V}$ and $I_1^{V}$, which has size $\leq 3$ by the definition of a Grassmann necklace. We have $n-1, n-2 \in I_{n-2}^{V} \setminus I_1^V$ for certain, and the third element (if it exists) must be $n$. On the other hand we have $n-5 \in I_1^{V} \setminus I_{n-2}^V$, along with at least one and at most two other elements.  If there is one such element call it $a$.  If there are two call them $a$ and $b$ with $a>b$.

  This means that the plusses in the Le diagram of $V$ coming from $I_{n-2}^{V}$ are as in the first part of Figure~\ref{fig messy}.  Stepping to $I_{n-1}^{V}$ simply removes the $(n-5)\rightarrow (n-2)$ path (second part of Figure~\ref{fig messy}), so $I_{n-1}^V$ contributes no new information about the Le diagram of $V$.

  \begin{figure}
    \includegraphics{messy}
    \caption{Plusses coming from $I_{n-2}^{V}$ (left) and  $I_{n-1}^{V}$ (right), in the case where $n-1, n-2\in I_{n-2}^{V}$.  The blue lines are the non-intersecting paths defining the position of the plusses.  The dashed blue lines may or may not appear, but if one appears then they both do.}\label{fig messy}
  \end{figure}

  \note{Remove the $I_n^V$ diagrams}

  %Stepping to $I_{n}^{V}$, $n-1$ is removed and either $n$ is put in if it was not there before, or one of $a$ or $b$ is put in and hence is no longer available as the starting point for a path.  This gives two possible configurations illustrated in the bottom two parts of Figure~\ref{fig messy}.

  With this information in hand, we can now tackle the points in the statement of the lemma. We know that 
  \[I_{n-2}^{W}  = (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\}, \quad n-5 \in I_1^W, \quad n-3\not\in I_{n-2}^W,\]
 and so the paths for building plusses from $I_{n-2}^{W}$ start at
 \[I_1^W \setminus I_{n-2}^W = \{n-3,n-5\} \cup (I_1^V \setminus I_n^V)\]
 and end at
 \[I_{n-2}^W \setminus I_1^W = \{n-2,n-1\}\cup(I_n^V \setminus I_1^V).\]
This means that $I_{n-2}^W$ contributes plusses as in Figure~\ref{fig messyD}, which proves point (1) and most of point (3). From equation \eqref{eq necklace} we see that $n-2$ appears only in $I_{n-2}^V$, so no other Grassmann necklace term for $V$ can contribute a plus in the $n-2$ column; this completes the proof of item (3).

\note{Modify this figure to include the case where there's no plus for column $n$.}

Now consider $I_{n-3}^{V}$.  By \eqref{eq necklace} $I_{n-3}^{V}$ contributes almost the same plusses as $I_{n-2}^{V}$: the only difference is that it contributes an $(n-5)\rightarrow (n-3)$ plus in place of the $(n-5)\rightarrow (n-2)$ plus.  Since we must have $n-3\in I_{n-3}^{V}$ (because the propagator $q$ exists in $V$), it follows from Lemma~\ref{lem I} that $I_{n-3}^{W} = I_{n-3}^{V}\cup\{n-2\}$.  Thus the paths for $I^{W}_{n-3}$ are the same as those for $I^{V}_{n-3}$ except that the path that did go to $n-3$ now goes to $n-2$.  This cannot conflict with another path since \eqref{eq necklace} shows that $n-2$ only appears in $I_{n-2}^{V}$ among the necklace elements of $V$.

This proves point (2), and combined with the observation above that $I_{n-1}^V$ contributes no new plusses compared to $I_{n-2}^V$, we also obtain item (4).

  Finally, if the Le diagram of $V$ did admit a $(n-5)\rightarrow (n-1)$ plus, it could only have been contributed by a necklace term that doesn't contain $n-5$.  By equation \eqref{eq necklace} the only terms that could have this property are $I_{n-2}^{V}$, $I_{n-3}^{V}$, and $I_{n-4}^{V}$, and the analysis above shows that $I_{n-2}^V$ and $I_{n-3}^V$ do not contribute a plus in this position.

  Recall that $n-5$ is the largest row index in the Le diagram of $V$, and $n-5 \in I_1^V \setminus I_{n-4}^V$ so $I_{n-4}^V$ does contribute a path starting at $n-5$. Since $n-4 \in I_{n-4}^V$ by the location of the propagator $q$, we must also have a path ending at $n-4$. Since the paths cannot cross, this implies that $I_{n-4}^V$ must contribute a $(n-5) \rightarrow (n-4)$ plus, and hence cannot contribute a $(n-5)\rightarrow (n-1)$ plus. This proves point (5) and completes the proof of the lemma.
\end{proof}
